---
title: "Lepr TRAP"
author: "Alan Rupp"
date: "`r format(Sys.time(), '%b %d, %Y')`"
output: pdf_document
geometry: margin=0.5in
---

Analyzing data from VMH-targeted TRAP-seq of LepRb-Cre mice looking for enriched genes.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE,
                      fig.align = "center")
library(tidyverse)
library(knitr); library(kableExtra)
```

```{r}
# read in count data
count_files <- list.files("../data/counts", full.names = TRUE)
samples <- str_extract(count_files, "Sample_[0-9]+")
counts <- map(
  count_files,
  ~ read_tsv(.x, skip = 4, col_names = FALSE)
)
counts <- bind_cols(counts) %>% 
  select(X1, starts_with("X2")) %>%
  set_names("gene_id", samples)
```

```{r}
# read in metadata
metadata <- read_csv("../data/Run_1607/Run_1607_mmyers.csv", skip = 18) %>%
  filter(!duplicated(Sample_ID)) %>%
  filter(Sample_ID %in% samples) %>%
  mutate("Cells" = ifelse(str_detect(Description, "LepR"), "Bead", "Sup")) %>%
  mutate(Cells = factor(Cells, levels = c("Sup", "Bead"))) %>%
  group_by(Cells) %>%
  mutate("Pair" = as.character(seq(n()))) %>%
  ungroup()
```

There are `r length(unique(metadata$Sample_ID))` samples. 

```{r}
metadata %>% select(Sample_ID, Description, Cells) %>%
  kable(caption = "Samples") %>%
  kable_styling(latex_options = "hold_position")
```

```{r}
# read in gene_id to gene_name mapper
tx2gene <- read_csv("../data/tx2gene.csv") %>% 
  filter(!duplicated(gene_id)) %>%
  select(starts_with("gene"))
```

# Differential expression
Using DESeq2 `r packageVersion("DESeq2")` to identify significantly enriched genes.

```{r}
dds <- DESeq2::DESeqDataSetFromMatrix(
  countData = counts %>% as.data.frame() %>% column_to_rownames("gene_id"),
  colData = metadata,
  design = ~ Pair + Cells
) %>% 
  DESeq2::DESeq()

enrichment <- DESeq2::results(dds) %>% as.data.frame() %>%
  rownames_to_column("gene_id") %>%
  left_join(., tx2gene, by = "gene_id") %>%
  select(gene_name, log2FoldChange, padj) %>%
  rename("Enrichment" = log2FoldChange, "P" = padj)

cpm <- DESeq2::fpm(dds) %>% as.data.frame() %>%
  rownames_to_column("gene_id") %>%
  left_join(., tx2gene, by = "gene_id") %>%
  select(-gene_id)

cpm_mean <- map(
  unique(metadata$Cells),
  ~ rowMeans(cpm[,filter(metadata, Cells == .x)$Sample_ID])
) %>%
  bind_cols() %>%
  set_names(unique(metadata$Cells)) %>%
  mutate("gene_name" = cpm$gene_name)
```

There are `r sum(enrichment$Enrichment > 0 & enrichment$P < 0.05, na.rm = TRUE)` significantly enriched genes.

## Top genes
```{r}
enrichment %>% arrange(desc(P < 0.05), desc(Enrichment)) %>%
  filter(!str_detect(gene_name, "^Gm")) %>%
  mutate(Enrichment = round(Enrichment, 2)) %>%
  mutate(P = format(P, digits = 2, scientific = TRUE)) %>%
  rename("Gene" = gene_name) %>%
  slice(1:20) %>%
  kable(caption = "Top 20 enriched genes") %>%
  kable_styling(latex_options = "hold_position")
```

---

I'm saving the DESeq2 output as `enrichment.csv`.

```{r}
# save file
if (!dir.exists("../results")) dir.create("../results")

enrichment %>% left_join(., cpm_mean, by = "gene_name") %>%
  rename("Gene" = gene_name) %>%
  select(Gene, Bead, Sup, Enrichment, P) %>%
  arrange(desc(P < 0.05), desc(Enrichment)) %>%
  mutate_at(vars(Bead, Sup, Enrichment), ~ round(.x, 2)) %>%
  mutate(P = format(P, digits = 2, scientific = TRUE)) %>%
  write.csv("../results/enrichment.csv", row.names = FALSE)
```

\pagebreak

# Session Info

```{r}
sessionInfo()
```